\documentclass[a4paper, 11pt]{article}
\usepackage{comment} % enables the use of multi-line comments (\ifx \fi) 
\usepackage{fullpage} % changes the margin
\usepackage[a4paper, total={7in, 10in}]{geometry}
\usepackage{amsmath,mathtools,mathdots}
\usepackage{amssymb,amsthm}  % assumes amsmath package installed
\usepackage{float}
\usepackage{xcolor}
\usepackage{mdframed}
\usepackage[shortlabels]{enumitem}
\usepackage{indentfirst}
\usepackage{hyperref}
\hypersetup{
	colorlinks=true,
	linkcolor=doc!80,
	citecolor=myr,
	filecolor=myr,      
	urlcolor=doc!80,
	pdftitle={Assignment}, %%%%%%%%%%%%%%%%   WRITE ASSIGNMENT PDF NAME  %%%%%%%%%%%%%%%%%%%%
}
\usepackage[most,many,breakable]{tcolorbox}
\usepackage{tikz}
\usepackage{caption}
\usepackage{kpfonts}
\usepackage{libertine}
\usepackage{physics}
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
\usepackage{mathrsfs}
\usepackage{tikz-cd}
\usepackage{float}

\definecolor{mytheorembg}{HTML}{F2F2F9}
\definecolor{mytheoremfr}{HTML}{00007B}
\definecolor{doc}{RGB}{0,60,110}
\definecolor{myg}{RGB}{56, 140, 70}
\definecolor{myb}{RGB}{45, 111, 177}
\definecolor{myr}{RGB}{199, 68, 64}

\input{../assignment-problem-box}

\newtheorem{lemma}{Lemma}
\renewenvironment{proof}{\noindent{\it \textbf{Proof:}}\hspace*{1em}}{\qed\bigskip\\}
% To give references for any problem use like this
% suppose the problem number is p3 then 2 options either 
% \hyperref[p:p3]{<text you want to use to hyperlink> \ref{p:p3}}
%                  or directly 
%                   \ref{p:p3}



\input{../../letterfonts}

\input{../../macros}

\setlength{\parindent}{0pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
	
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	
	\textsf{\noindent \large\textbf{Soham Chatterjee} \hfill \textbf{Assignment - 1}\\
		Email: \href{sohamc@cmi.ac.in}{sohamc@cmi.ac.in} \hfill Roll: BMC202175\\
		\normalsize Course: Algebra and Computation \hfill Date: \today}
	
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Problem 1
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	
\begin{problem}{%problem statement
		Problem Set 1: P5
	}{p1% problem reference text
}
For a prime $p$ and a positive integer $e$, prove that $\bbZ_{p^e}^*$ is cyclic.
\end{problem}
\solve{
	We will prove this in 3 stages: $e=1$, $e=2$, $e> 2$.
	
	\section*{Case 1: $e=1$}
	\begin{lemma}\label{totprop}
		$\sum\limits_{d\mid n} \varphi(d)=n$
	\end{lemma}
	\begin{proof}
		Consider the list of numbers $S=\lt\{\frac1n,\frac2n,\dots,\frac{n}{n}\rt\}$. If we express every number in $S$ as simplified form i.e. $\frac{p}{q}$ form where $gcd(p,q)=1$. Then the denominators are all the divisors of $n$. 
		
		
		Then for any $k\in[n]$ we have $$\frac{k}{n}=\frac{\frac{k}{gcd(k,n)}}{\frac{n}{gcd(k,n)}}$$Denote $d_k\coloneqq\frac{n}{gcd(k,n)}$ then $d_k$ is a factor of $n$. And since $gcd\lt( \frac{k}{gcd(k,n)}, \frac{n}{gcd(k,n)} \rt)=1$ we have $\frac{k}{gcd(k,n)}\in \bbZ_{d_k}^*$. Let $k\in\bbZ_{d}^*$ then suppose $l$ is such that $d\times l=n$ then the fraction $\frac{k}{d}=\frac{k\times l}{n}\in S$ and its simplified form is infact $\frac{k}{d}$.
		
		Hence for any $d\mid n$, the number of fractions with denominator $d$ is $\varphi(d)$, since for all such fractions the numerators are the elements of $\bbZ_{d}^*$. Therefore we have $\sum\limits_{d\mid n} \varphi(d)=n$.
	\end{proof}
	
	Now define for $d$ such that $d\mid p-1$, $S_d=\{a\in \bbZ_p^*\mid ord(a)=d\}$. Then we have the following lemma:
	
	\begin{lemma}
		$|S_d|=\varphi(d)$
	\end{lemma}
	\begin{proof}
		First we will show that $|S_d|\in\{0,\varphi(d)\}$ then we will show that $|S_d|=\varphi(d)$. Now if $|S_d|\neq 0$ then $\exs\ a\in S_d$ such that $ord(a)=d$. Then consider the polynomial $x^d-1$ over $\bbF_p$. $1,a,a^2,\dots,a^{p-1}$ are its distinct roots. Since the degree is $d$ these are the only roots of the polynomial. Now $a^k$ has order $\frac{d}{gcd(d,k)}$. Then the elements which has order $d$ are $a^k$ where $gcd(k,d)=1$. Hence there are $\varphi(d)$ many powers of $a$ which has order $d$. Therefore $|S_d|\in\{0,\varphi(d)\}$.
		
		Now we have by \lmref{totprop} $$\sum_{d\mid p-1}\varphi(d)=p-1$$Now $\{S_d\colon d\mid p-1\}$ is a partition of $\bbZ_P^*$. Therefore $\sum\limits_{d\mid p-1}|S_d|=p-1$. Hence $$p-1=\sum\limits_{d\mid p-1}|S_d|\leq \sum\limits_{d\mid p-1}\varphi(d)=p-1\iff |S_d|=\varphi(d)\ \forall\ d\text{ such that } d\mid p-1$$
	\end{proof}
	
	Hence the number of elements in $\bbZ_p^*$ which has order $d$ such that $d\mid p-1$
	
	Now we will introduce another definition. Let $H$ be a group. Then Exponent of $H$ is the smallest number $n$ such that $\forall a\in H$, $a^n=1$. Now we will show that every finite abelian group has an element which has the order to be exponent of the group. Then we will show that $\bbZ_p^*$ has exponent $p-1$. With that we can say $\bbZ_p^*$ has an element which has order $p-1$. Therefore $\bbZ_p^*$ is cyclic since $|\bbZ_p^*|=p-1$ because $\bbZ_p^*$ is a finite abelian group.
	\begin{lemma}
		If $G$ is a finite abelian group with exponent $n$ then $\exs\ g\in G$ such that $ord(g)=n$.
	\end{lemma}
	\begin{proof}
		By structure theorem we have $$G\cong \bbZ_{q_1}\times \cdots \times \bbZ_{q_m}$$ where $q_1,\dots, q_m$ are primes powers. Now $\forall\ g\in G$, $ord(g)\mid lcm(q_1,\dots,q_m)$. The element in $\bbZ_{q_1}\times \cdots \times \bbZ_{q_m}$, $(1,1,\dots, 1)$ has order $lcm(q_1,\dots,q_m)$. So the exponent of $G$ is $lcm(q_1,\dots,q_m)$ and the corresponding element of $(1,\dots,1)$ has order $lcm(q_1,\dots,q_m)$.
	\end{proof}
	\begin{lemma}
		$\bbZ_p^*$ has exponent $p-1$.
	\end{lemma}
	\begin{proof}
		Over $\bbF_p$ the equation $x^{p-1}-1$ has $p-1$ roots which are all the elements of $\bbZ_p^*$. There does not exists any polynomial of lower degree which satisfies this property. Hence the exponent of $\bbZ_p^*$ is $p-1$.
	\end{proof}
	
	Therefore there exists an element of $\bbZ_p^*$ which has order $p-1$. Therefore the group $\bbZ_p^*$ is cyclic.
	\section*{Case 2: $e=2$}
	\begin{lemma}
		Let $g$ be generator of the group $\bbZ_p^*$. Then either $g$ or $g+p$ is generator for $\bbZ_{p^2}^*$.
	\end{lemma}
	
	\begin{proof}
		We have $|\bbZ_{p^2}^*|\varphi(p^2)=p(p-1)$. Let $g$ has order $m$ in $\bbZ_{p^2}^*$. Then $g^p\equiv 1\bmod p$. Hence $p-1\mid m$. Therefore $m=p(p-1)$ or $m=p-1$ since $m\mid p(p-1)$. If its the first case then we are done. For the later take the element $g+p$. Again let its order is $m'$. Then $(g+p)^{m'}\equiv 1\bmod p$. So $p-1\mid m'$. Hence $m'$ can be either $p-1$ or $p(p-1)$. If it is also $p-1$ then we have \begin{align*}
			1\equiv (g+p)^{p-1} & \equiv g^{p-1}+(p-1)g^{p-2}p+p^2(\cdots)\bmod {p^2}\\
			& \equiv g^{p-1}+p(p-1)g^{p-2}\bmod {p^2}\\
			& \equiv 1+p(p-1)g^{p-2}\bmod {p^2}
		\end{align*}
		Therefore $$p(p-1)g^{p-2}\equiv 0\bmod{p^2}\iff p\mid (p-1)g^{p-2}$$ which is not possible since $gcd(p,p-1)=1$ and $gcd(p,g)=1$. Contradiction. Hence at least one of $g$ and $g+p$ has order $p(p-1)$.  
	\end{proof}
	
	With this lemma we have an element of $\bbZ_{p^2}^*$ which has order $p(p-1)=|\bbZ_{p^2}^*|$. So $\bbZ_{p^2}^*$ is cyclic.
	
	
	
	
	\section*{Case 3: $e> 2$}
	\begin{lemma}
		$(1+p)^{p^k}\equiv 1+p^{k+1}\bmod {p^{k+2}}$
	\end{lemma}
	\begin{proof}
		\begin{align*}
			(1+p)^{p^k}& \equiv \lt((1+p)^p   \rt)^{p^{k-1}}\\
			& \equiv \lt( 1+p^2+\binom{p}{2}p^2 \rt)^{p^{k-1}}\bmod {p^{k+2}}\\
			& \equiv 1+p^2\times p^{k-1}\bmod p^{k+2}\\
			&\equiv 1+p^{k+1}\bmod{p^{k+2}}
		\end{align*}
	\end{proof}
	Therefore $$(1+p)^{p^{k+1}}\equiv(1+p^{k+1})^{p}\equiv 1+p\times p^{k+1}\equiv 1+p^{k+2}\equiv 1\bmod{p^{k+2}}$$Hence $(1+p)$ has order $p^{k+1}$ in $\bbZ_{p^{k+2}}^*$. Or we can say $1+p$ has order $p^{e-1}$ is $\bbZ_{p^e}^*$. 
	
	Let $g$ be the generator of $\bbZ_p^*$. Then let the order of $g$ in $\bbZ_{p^e}^*$ is $m$. Then $p-1\mid m$. So $g$ has order $p^k(p-1)$. Therefore the number $g(1+p)\bmod {p^e}$ has order $p^{e-1}(1-p)=\varphi(p^e)$. Therefore $\bbZ_{p^e}^*$ is a cyclic group.
	
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Problem 2
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{problem}{%problem statement
		Problem Set 1: P6
	}{p2% problem reference text
	}
Let $N=p_1p_2\cdots p_k$ be a Carmichael number and $p_i$'s are primes. In class we have seen that given $N$ as input, a single pass of Miller-Robin primality test outputs a nontrivial factor of $N$ with probability $\geq \frac12$. We can do a finer calculation and get better success probability. Show that a single pass of Miller-Robin primality test outputs a nontrivial factor of $N$ with probability $1-\frac1{2^{k-1}}$.
\end{problem}
\solve{
	Let $\phi$ be the isomorphism of $$\bbZ_N^*\cong \bbZ_{p_1}^*\times \cdots \times \bbZ_{p_k}^*$$Now suppose $N-1=2^vm$ where $m$ is odd. Let $a\in \{2,\dots, N-2\}$ Let $l_a$ be the minimum such that  $a^{2^{l_a+1}m} \bmod N\equiv 1$. Surely for all $a$, $l_a>0$ and $l_a\leq N-1$ Now take $l=\max\{l_a\mid a\in \{2,\dots, N-2\}\}$. Therefore $l>0$ and $l\leq N-1$. For all $k<l$ there exists $a\in\{2,\dots, N-2\}$ such that  $a^{2^{k+1}m}\not\equiv 1\bmod N$. 
	
	Now consider the group $$G_N=\{a\in \bbZ_N^*\mid a^{2^lm}\equiv \pm 1\bmod N\}$$Now there exists at least one $\tda$ such that $\tda^{2^lm}\equiv -1\bmod N$ since otherwise for all $a\in \{2,\dots N-2\}$, $l_a\leq l-1$. Then $\max\{l_a\mid a\in\{2,\dots ,N-2\}\}\leq l-1$ which contradicts that the value we got is $l$. Hence there exist a $\tda\in \bbZ_N^*$ such that $\tda^{2^lm}\equiv -1\bmod N$.
	
	Now $\phi(\tda^{2^lm})=(-1,\dots,-1)$. Suppose $\phi(\tda)=(\tda_1,\dots, \tda_k)$. Then we have $$\forall\ i\in[k],\ \tda_i^{2^lm}\equiv -1\bmod p_i$$Now for any $i\in[k]$ the corresponding element in $\bbZ_N^*$ of $(1,\dots,1,\tda_i,1,\dots,1)$ denote by $g$. Then $g^{2^lm}\not\equiv -1\bmod N$. There are $k$ many slots here and in each slot we have 2 options $1$ or $\tda_i$. Hence with above like construction we can have at most $2^k$ many elements. Among these the elements $(1,\dots, 1)$ and $(\tda_1,\dots,\tda_k)$ are in $G_N$. All the other elements remain in distict cosets of $G_N$ in $\bbZ_N^*/G_N$. Hence $$Pr_{a\in_R \bbZ_N^*}[a\in \bbZ_N^*-G_N]\geq \frac{2^k-2}{2^k}=1-\frac1{2^{k-1}}$$Hence $$Pr[\text{Primality Test outputs a nontrivial factor of $N$}]\geq 1-\frac1{2^{k-1}}$$
	
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Problem 3
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{problem}{%problem statement
		Problem Set 1: P7
	}{p3% problem reference text
	}
	Design a randomized polynomial time algorithm such that given $N$ and $\varphi(N)$ as inputs, it outputs a nontrivial factor of $N$ with probability at least $\frac12$, where $\varphi(\cdot)$ is the Euler’s totient function
\end{problem}
\solve{
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Problem 4
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{problem}{%problem statement
		Problem Set 1: P13
	}{p4% problem reference text
	}
	Design a deterministic polynomial time algorithm to compute the gcd of two univariate polynomials
	using resultants and linear system solving.
\end{problem}
\solve{
	Let $p,q\in \bbF[x]$ where $\deg p=m$ and $\deg q=n$. The Sylvester matrix generated by $p,q$ is $S_{p,q}$. Let for any $k\in\bbN$, $\bbF_k\coloneqq \{f\in \bbF[x]\mid \deg f<k\}$. Then for $(u,v)\in\bbF_n\times \bbF_m$, $S_{p,q}(u,v)=up+vq$.
	
	Let $gcd(p,q)=h$ and $\deg h=d$.
	\begin{lemma}
		$\dim \ker S_{p,q}=\deg gcd(p,q)$
	\end{lemma}
	\begin{proof}
		Let $(x,y)\in \ker S_{p,q}$. Then $px+qy=0$. Now denote $p=hp_0$ and $q=hq_0$. Hence $gcd(p_0,q_0)=1$. Therefore $$px+qy=0\iff p_0x+q_0y=0\iff p_0x=-q_0y$$Therefore $q_0\mid x$ and $p_0\mid y$. So let $x=q_0g_x$ and $y=p_0g_y$. Then $$p_0x+q_0y=0\iff p_0q_0g_x+q_0p_0g_y=0\iff p_0q_0(g_x+g_y)=0\iff g_x=-g_y$$
		So denote $g=g_x=-g_y$. So $x=q_0g$, $y=-p_0g$. Now  $$\deg x<\deg q\iff \deg q_0+\deg g<\deg q_0+\deg h\iff \deg g<\deg h $$  Hence for each $(x,y)\in \ket S_{p,q}$ there exists unique $g\in \bbF_d$ such that $x=q_0g$ and $y=-p_0g$ and also for each $g\in \bbF_d$ we have $x=q_0g$ and $y=-p_0g$ such that $px+qy=0$. Hence there exists a bijection $\bbF_d \cong \ker S_{p,q}$ by $g\mapsto (q_0g,-p_0g)$
	\end{proof}
	
	Therefore by Rank-Nullity Theorem $$\rank(S_{p,q})+\dim \ker S_{p,q}=m+n$$ Therefore $\rank(S_{p,q})=m+n-d$. Hence the last $d$ rows of the row echelon form of the $S_{p,q}^T$ are zeros. Let $(S_{p,q}^T)^*$ denote the row echelon form of $S^T_{p,q}$. Let $e_i$ denote the $i$th row of $(S_{p,q}^T)^*$. Hence the last nonzero row of $(S^T_{p,q})^*$ is $e_{m+n-d}$. We have $\deg e_{m+n-d}\leq d$. Now for $i\in[n]$ the $i$th row of $S^T_{p,q}$ is just $x^{n-i}p$ and for $n+1\leq j\leq n+m$ the $j$th row is $x^{m+n-j}q$. Hence $$e_{m+n-d}=\sum_{i=1}^{n}\alpha_i x^{n-i}p+\sum_{i=n+1}^{m+n}\alpha_ix^{m+n-i}q$$ The LHS has degree $\leq d$ and the RHS is divisible by $h$ since $h\mid p$ and $h\mid q$. Hence $h=e_{m+n-d}$ up to some unit multiplication. Therefore we can say $e_{m+n-d}$ is the gcd of $p,q$. Therefore the algorithm will be 
	\parinf
	
	\textbf{Algorithm:}\begin{enumerate}[label=Step \arabic*,leftmargin=1.5cm]
		\item Construct $S_{p,q}$
		\item Find Row Echelon Form of $S_{p,q}^T$ by Gaussian Elimination
		\item Output the last nonzero row 
	\end{enumerate}
	\parinn
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Problem 5
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{problem}{%problem statement
		Problem Set 1: P14
	}{p5% problem reference text
	}
	Give a polynomial time algorithm to compute the gcd of two bivariate polynomials, without using
	bivariate factorization.
\end{problem}
\solve{
		\begin{lemma}\label{gcdlc}
		Let $R$ be an Euclidean Domain. Let $p\in R$ be a prime and $f,g\in R[x]$ be nonzero. Let $h=gcd(f,g)\in R[x]$. Denote $\ov{f}=f\bmod p$ and $\ov{g}=g\bmod p$ and $d=\deg h$ and $\alpha=lc(h)$. Assume $p\nmid b=gcd(lc(f),lc(g))\in R$ and $\ovd=\deg gcd(\ovf,\ovg)$. Then \begin{enumerate}
			\item $\alpha\mid b$
			\item $\ovd\geq d$
			\item $d=\ovd\iff \ov{\alpha}\cdot gcd(\ovf,\ovg)=\ovh\iff p\nmid \res\lt(\frac{f}{h},\frac{g}{h}\rt) $
		\end{enumerate}
	\end{lemma}
	
	
	\begin{proof}
		\begin{enumerate}
			\item 	Now $h$ divides both $f,g$. Therefore $lc(h)$ divides both $lc(f)$ and $lc(g)$ in $R$. Hence $\alpha\mid b$
			\item Let $u=\frac{f}{h}$ and $v=\frac{g}{h}$. Since $p\nmid b\implies p\nmid lc(h)$. Hence $\deg h=\deg \ovh=d$. Now $$\ovu\ovh=\ovf\text{ and }\ovv\ovh=\ovg$$ Hence $\ovh\mid \ovf$ and $\ovh\mid \ovg\implies \ovh\mid gcd(\ovf,\ovg)$. Therefore $\deg gcd(\ovf,\ovg)\geq \deg \ovh\implies \ovd\geq d$.
			\item $d=\ovd\iff \deg \ovh=\deg gcd(\ovf,\ovg)$. Now $p\nmid b$ and $\alpha\mid b$ so $p\nmid \alpha$. Hence $\alpha$ is a unit in $R/\la p\ra$ as $R/\la p\ra$ is a field. In a field gcd is always taken to be monic. Now $\ov{\alpha}=lc(\ovh)$. Since $\deg \ovh=\deg gcd(\ovf,\ovg)$ we can say $\ovh=u\cdot gcd(\ovf,\ovg)$ for some unit $u\in R/\la p\ra$. Now since $gcd(\ovf,\ovg)$ is monic we have $u=\ov{\alpha}$ Therefore $d=\ovd\implies \ov{\alpha}\cdot gcd(\ovf,\ovg)=\ovh$. Other direction obviously becomes true as $\ov{\alpha}$ is a unit in $R/\la p\ra$. 
			
			Now $p\nmid b\implies p$ divides at most one of $lc(u)$ or $lc(v)$. WLOG assume $p\nmid lc(u)$. We know $$p\mid \res (u,v)\iff gcd(\ovu,\ovv)\neq 1\text{ in }R/\la p\ra$$So \begin{align*}
				gcd(\ovf,\ovg)=gcd(\ovu,\ovv)\cdot \frac{\ovh}{\ov{\alpha}} & \iff \ov{\alpha} gcd(\ovf,\ovg)=gcd(\ovu,\ovv)\ovh \\
				                                                            & \iff \ovh=gcd(\ovu,\ovv)\ovh                       \\
				                                                            & \iff gcd(\ovu,\ovv)=1                              \\
				                                                            & \iff p\nmid \res(\ovu,\ovv)                        \\
				                                                            & \iff p\nmid \res(u,v)
			\end{align*}
			
		\end{enumerate}
	\end{proof}
	

	
	\begin{algorithm}[H]
		\KwIn{\begin{enumerate}
				\item Primitive Polynomials $f,g\in\bbF[x,y]=R[x]$
				\item $\deg_x f=n\geq \deg_x g\geq 1$
				\item $\deg_y f,\deg _yg\leq d$
		\end{enumerate}}
	\KwOut{$h=gcd(f,g)\in \bbF[x,y]$}
	\DontPrintSemicolon
		\Begin{
		$b\longleftarrow gcd(lc(f),lc(g))$, $FAIL\longleftarrow 1$\;
		\While{$FAIL$}{
			Choose a random monic irreducible polynomial $p\in \bbF[y]$ with $\deg p=d+1+\deg b$\;
			$\ov{f}\longleftarrow f\bmod p$, $\ov{g}\longleftarrow g\bmod p$\;
			Use Extended Euclidean Algorithm over $\bbF[y]/\la p\ra$ on $\ov{f}$ and $\ov{g}$ to compute the monic $v\in R/\la p\ra$\;
			Compute $w,f',g'\in R[x]$ with $\deg_yw,\deg_yf',\deg_yg'<\deg p$ such that:$$w\equiv bv\bmod p\qquad f'w\equiv bf\bmod p\qquad g'w\equiv bg \bmod p$$\;
			\If{$\deg_y(f'w)=\deg_y(bf)$ and $\deg_y(g'w)=\deg(bg)$}{$FAIL\longleftarrow 0$}
			\Return{ premitive part of $w$ w.r.t $x$}
		}
		}
		\caption{Modular Bivariate GCD Algorithm}
	\end{algorithm}
	
		Now in $\bbF[x,y]$ let $gcd(f,g)=h$ and $r=\res_x\lt(\frac{f}{h},\frac{g}{h}\rt)\in \bbF[y]$. Now $\deg_yb<\deg_yp=\deg p$ and hence $p\nmid b$. Assume $p\nmid r$ then by \lmref{gcdlc} we have $\alpha \cdot v \equiv h\bmod p$ and $\alpha\mid b$. Therefore $$w\equiv bv\equiv \lt(\frac{b}{\alpha}\rt)h\bmod p$$ Now primitive part of $w$=premitive part of $\lt(\frac{b}{\alpha}\rt)h$=$h$. Hence correct result is returned.
		
				Now if $p\mid r$ then by \lmref{gcdlc} we have $\deg_xgcd(\ovf,\ovg)>\deg_x h$. If the degree conditions in step 8 are satisfied then the congruences in step 7 would be equalities and the primitive part of $w$ will be a common divisor of $f$ and $g$ of higher degree than $\deg_x h$. Contradiction. So the degree condtions will not be satisfied.
 }
\end{document}
