\chapter{Potential Games}
\section{Best Response Dynamics}
The existence of a Nash equilibrium is clearly a desirable property of a strategic game. In this chapter and the next we discuss some natural classes of games that do have a Nash equilibrium. The \textit{Best-Response-Dynamics} is a straightforward procedure by which players search for a pure
Nash equilibrium ($\pne$) of a game. 

\begin{algorithm}\DontPrintSemicolon
\Begin{
\For{$t=1,\dots, T$}{
	\If{$t=1$}{Each player plays an arbitrary pure strategy}
	\Else{Pick a player $i\in[n]$\;
	$s_i^t\longleftarrow \arg\min\limits_{s_i\in S_i}c_i(s_i,s_{-i}^{t-1})$\;
$s_j^t\longleftarrow s_j^{t-1}\ \forall\ j\in[n]$, $j\neq i$}
}
}
	
\caption{\textsc{Best-Response-Dynamics} (\textsf{BRD})}
\end{algorithm}


\nt{Best-response dynamics can only halt at a $\pne$ and it cycles in any game without one. It can also cycle in games that have a $\pne$. For example consider the following 2 player.}

\section{Network (Atomic) Congestion Games}
\begin{definition}{Network (Atomic) Congestion Games}{}
	A network (atomic) congestion game or in short NCG consists of the following:
	\begin{itemize}[itemsep=-1mm]
		\item A directed graph $G=(V,E)$.
		\item $N$ players where  each player $i\in[n]$ has some source-sink pair $(s_i,t_i)\in V\times V$ associated with it.
		\item Edge cost functions $c_e:[n]\to \bbR$ for each edge $e\in E$.
		\item Player $i\in[N]$ has strategy set $S_i=$ Set of all $s_i\rightsquigarrow t_i$ paths in $G$. $S=\bigtimes\limits_{i=1}^N S_i$.
		\item For a strategy profile $f\in S$ (often called \textit{flow}), let $n_e(f)=|\{i\colon e\in f_i\}|$. Then the cost to player $i$ of strategy profile $f$ is $C_i(f)=\sum\limits_{e\in s_i}c_e(n_e(f))$. 
	\end{itemize}
\end{definition}

So we can define (atomic) NCG by the tuple $$\Big(G=(V,E),N,\{(s_i,t_i)\mid i\in[N]\}, \{c_e:[N]\to \bbR_{\geq 0}\mid e\in E\}\Big)$$

Note that unlike the last few lectures where we’ve been talking about utility-maximization games,
this is a \hyperref[def:cost-min-game]{cost-minimization game}. But of course we could just let a player’s utility be the negative
of its cost and everything would work as you expect.
\begin{lemma}{}{ncg-pne}
	Every NCG has a $\pne$.
\end{lemma}
\begin{proof}
	Given a strategy profile $f\in S$, we will define a potential function $\Phi:S\to\bbR_{\geq 0}$ with the property that if $f$ is not an equilibrium then $\exs\ f'\in S$ such that $\Phi(f)>\Phi(f')$. Thus if $f^*\in S$ minimizes $\Phi$ then $f^*$ must be a $\pne$.
	
	Consider the potential function $\Phi:S\to\bbR_{\geq 0}$: $$\Phi(s)=\sum_{e\in E}\sum_{i=1}^{n_e(f)}c_e(i)$$Now it is enough to calculate the change in potential when a player deviates to any other strategy since for $f,f'\in S$ $$\Phi(f)-\Phi(f')=\sum_{i=0}^{N-1} \Phi(f^{(i)})-\Phi(f^{(i+1)})$$ where $f^{(i)}=(f_1',f_2',\dots, f_i',f_{i+1},\dots, f_N)$ and for $f^{(0)}=f$. Now for any strategy profile $f\in S$ if the player $i$ deviates to the strategy $f_i'\in S_i$ then \begin{align*}
		C_i(f)-C_i(f'_i,f_{-i}) & = \lt[  \sum_{e\in f_i\cap f_i'}c_e(n_e(f))+\sum_{e\in f_i\setminus f_i'}c_e(n_e(f))\rt]-\lt[  \sum_{e\in f_i\cap f_i'}c_e(n_e(f_i',f_{-i}))+\sum_{e\in f_i'\setminus f_i}c_e(n_e(f_i',f_{-i}))\rt]\\
		&=  \sum_{e\in f_i\cap f_i'}\underbrace{c_e(n_e(f))-c_e(n_e(f_i',f_{-i}))}_{=0} + \sum_{e\in f_i\setminus f_i'}c_e(n_e(f))-\sum_{e\in f_i'\setminus f_i}c_e(n_e(f_i',f_{-i}))\\
		& = \sum_{e\in f_i\setminus f_i'}c_e(n_e(f))-\sum_{e\in f_i'\setminus f_i}c_e(n_e(f)+1)\\
	\end{align*}Therefore the change in the potential is \begin{align*}
	\Phi(f)-\Phi(f_i',f_{-i}) & = \sum_{e\in E}\sum_{i=1}^{n_e(f)} c_e(i)-\sum_{e\in E}\sum_{i=1}^{n_e(f_i',f_{-i})} c_e(i)\\
	& =\sum_{e\in E}\lt[\sum_{i=1}^{n_e(f)} c_e(i)-\sum_{i=1}^{n_e(f_i',f_{-i})} c_e(i)\rt]\\
	&=\sum_{e\in f_i\setminus f_i'}c_e(n_e(f))-\sum_{e\in f_i'\setminus f_i}c_e(n_e(f)+1)\\
	& = C_i(f)-C_i(f'_i,f_{-i})
\end{align*}So the change in potential is exactly equal to the change in the cost of the player who deviates. Therefore if $f$ is not a $\pne$ then $\exs\ i\in[N]$ such that $\exs\ f_i'\in S_i$ such that $c_i(f)-c_i(f_i',f_{-i})>0$ and therefore $\Phi(f)-\Phi(f_i',f_{-i})>0$. Hence every NCG has a $\pne$.
\end{proof}

\section{Potential Games}
\begin{definition}{Potential Game}{}
	A game $\Gm$ is a potential game if there exists a potential function $\Phi:S\to\bbR_{\geq 0}$ where $S$ is the set of strategy profiles such that $\forall\ s\in S$ and $s_i'\in S_i$ $C_i(s)-C_i(s_i',s_{-i})=\Phi(s)-\Phi(s_i',s_{-i})$
\end{definition}
 
 In the proof of \thmref{ncg-pne} we showed that every NCG is a potential game. Now we will show that every potential game has a $\pne$.
 \begin{Theorem}{}{}
 	Every potential game has a Pure Nash Equilibrium
 \end{Theorem}
\begin{proof}
	For a potential game $\Gm$ let $\Phi$ is the potential function for $\Gm$. Then $C_i(s)-C_i(s_i',s_{-i})=\Phi(s)-\Phi(s_i',s_{-i})$. Now consider the strategy profile $s=\arg\min\limits_{s\in S}\Phi(s)$. If any player had incentive to deviate there would be a strategy profile with smaller potential which is not possible by the definition of $s$. Therefore $s$ also has the minimum cost. Therefore $s$ is $\pne$. 
\end{proof}

\begin{lemma}{}{}
	Best Response Dynamics cannot cycle in a potential game.
\end{lemma}
\begin{proof}
	In each iteration of the $\brd$ every time any player deviates to play a best response the potential must decrease. Hence $\brd$ cannot cycle.
\end{proof}

Suppose there exists a time $T$ such that every player was chosen in the $\brd$ to choose their best response in the Best response algorithm. Then:
\begin{lemma}{}{}
	Let $s^*\in S$ be the strategy profile at time $t$. If $s^*$ is the strategy profile after $T$ further steps of $\brd$ then $s^*$ is a $\pne$.
\end{lemma}
\begin{proof}
	Since in every $T$ steps every player has the option to deviate to another strategy but chose not to. Therefore for each player $i\in[N]$, for all $s_i'\in S_i$, $C_i(s)\leq C_i(s_i',s_{-i})$. Therefore clearly $s^*$ is a $\pne$. 
\end{proof}
\begin{lemma}{}{}
	Let $s^\in S$ be the strategy profile after $T|S|$ steps of $\brd$. Then $s^*$ is a $\pne$.
\end{lemma}
\begin{proof}
	Since $\brd$ cannot cycle, $\exs\ s\in S$ that must have persisted fro $T$ time steps. Therefore by the previous lemma this must be a $\pne$.
\end{proof}

\begin{Theorem}{}{}
	In a finite potential game from an arbitrary initial outcome the Best Response Dynamics converges to a $\pne$ if $\exs\ T\in\bbN$ such that in every $T$ steps of $\brd$ every player is chosen at least once.
\end{Theorem}

Since every (Atomic) NCG is a potential game we have the following corollary:
\begin{corolary}{}{}
	In an (Atomic) NCG, $\brd$ converges to a $\pne$ if $\exs\ T\in\bbN$ such that in every $T$ steps of $\brd$ every player is chosen at least once. or ``every player is chosen infinitely often".
\end{corolary}


\subsection{General Congestion Games}
General Congestion Games are generalized version of (atomic) NCG. We will show that they are also potential game.
\begin{definition}{General Congestion Games}{}
	A basic definition general Congestion Games or CG consists of the following: $$\Big(E,N, \{S_i\mid i\in[N]\}, \{c_e:[N]\to \bbR_{\geq 0}\mid e\in E\}\Big)$$
	\begin{itemize}[itemsep=-1mm]
		\item A base set $E$ of congestible elements.
		\item $N$ players.
		\item For each player $i\in[N]$ a finite set of strategies $S_i$ where $S_i\subseteq 2^E$. $S=\bigtimes\limits_{i=1}^N S_i$.
		\item Cost functions $c_e:[N]\to \bbR$ for each element $e\in E$.
		\item For a strategy profile $s\in S$ (often called \textit{flow}), let $n_e(s)=|\{i\colon e\in s_i\}|$. Then the cost to player $i$ of strategy profile $s$ is $C_i(s)=\sum\limits_{e\in s_i}c_e(n_e(s))$. 
	\end{itemize}
\end{definition}

 Consider the  function $\Phi:S\to \bbR_{\geq 0}$ where for any strategy profile $s\in S$, $$\Phi(s)=\sum\limits_{e\in E}\sum\limits_{i=1}^{n_e(s)}c_e(i)$$ that is the same function as the potential function in the case of NCG. This is also a potential function for general CG's which makes general CG's are also potential game.
\subsection{Max Cut Game}

\begin{definition}{Max Cut Game}{}
	A max cut game consists of the following:
	\begin{enumerate}[itemsep=-1mm]
		\item An undirected weighted graph, $G=(V,E)$ and $w:E\to\bbR$.
		\item $N$ players.
		\item For each player $i\in[N]$, has 2 strategies: $S_i=\{L,R\}$. $S=\bigtimes\limits_{i=1}^N S_i$.	
		\item Utility functions $u_i:S\to \bbR_{\geq 0}$ for each player $i\in[N]$. For any strategy profile $s\in S$,  $u_i(s)=\sum\limits_{\substack{e=\{i,j\}\\ s_i\neq s_j}}w_e$
	\end{enumerate}
\end{definition}

So like general congestion games we can denote a Max Cut game by the tuple $(G,w,N)$. The max cut game is also a potential game. Consider the potential function $\Phi:S\to\bbR_{\geq 0}$ where for any strategy profile $s\in S$, $$\Phi(s)=\sum_{\substack{e=\{i,j\}\\ s)i\neq s_j}}w_e$$With this function we can prove that the Max Cut game is indeed a potential game and henceforth there exists a $\pne$.

\section{Class: \textsf{PLS}}
\begin{definition}{$\pls$ (Polynomial Local Search)}{}
	A local search problem $L$ has a set of problem instances $D_L\subseteq \Sg^*$ where any $I\in D_L$ is a particular problem instance. For each instance $I\in D_L$ there exists a finite solution set $F_L(I)\subseteq \Sg^*$. Let $R_L$ be the relation that models $L$ i.e. $$R_L\coloneqq \{(I,s)\mid I\in D_L,s\in F_L(I)\}$$ Then $R_L$ is in $\pls$ if:\begin{enumerate}[label=(\roman*)]
		\item The size of every solution $s\in F_L(I)$ for any $I\in D_L$ is polynomially bounded in the size of $I$.
		\item The problem instances $I\in D_L$ and the solutions $s\in F_L(I)$ are polynomial time verifiable.
		\item There is a polynomial time computable function $C_L:\Sg^*\times \Sg^*\to \bbR_{\geq 0}$ that returns for each $I\in D_L$ and each $s\in F_L(I)$ the cost $C_L(I,s)$.
		\item There is a polynomial time computable function $N:(I,s)\mapsto S$ where $S\subseteq F_L(I)$ i.e. returns the set of neighbors of lower cost for each $I\in D_L$ and each $s\in F_L(I)$ or states the $s$ is locally optimal.
	\end{enumerate}
\end{definition}

Note that for each $I\in D_L$ and each $s\in F_L(I)$ using (iii) and (iv) we can find a neighboring solutions of lower cost of $s$ or determine $s$ is locally minimal. The problem we want to focus is to find a locally minimal cost solution given an instance $I$ of $L$.

\begin{goal}
	Find a locally minimal cost solution given an instance $I$ of $L$.
\end{goal}
Finding $\pne$ in a Congestion Game (\textsf{PNE-CG}) is in $\pls$. Also finding $\pne $  in a Max Cut Game is in $\pls$.
\begin{lemma}{}{}
	$\textsf{PNE-CG}\in \pls$
\end{lemma}
\begin{proof}
	We will show that computing $\pne$ in congestion games is an instance of $\pls$ by fist describing conditions needed for $\pls$. Suppose $\Big(E,N, \{S_i\mid i\in[N]\}, \{c_e:[N]\to \bbR_{\geq 0}\mid e\in E\}\Big)$ be any instance of CG.
	\begin{enumerate}[label=(\roman*)]
		\item Any strategy profile is a solution for a specific instance of congestion games. Now each element of strategy profile by definition of CG is a subset of the base set $E$. Therefore size of any strategy profile is polynomially bounded by the size of the instance of the CG.
		\item For solution for the CG has to be valid strategy profile. So given any $N$-tuple of subsets of $E$ we can check if the $i^{th}$ element of the tuple is in $S_i$ for all $i\in[N]$ in polynomial time.
		\item For any strategy profile $s\in S$ we return the potential function for the CG as the cost function $\Phi(s)=\sum\limits_{e\in E}\sum\limits_{j=1}^{n_e(s)}c_e(i)$. This function is a polynomial time computable function.
		\item Given a strategy profile $s$ it checks if it is a $\pne$ by checking if we can switch the strategy of agent $i$ from $s_i$ to $s_i'$ such that $\Phi(s)>\Phi(s_i',s_{-i})$ for all agents.  If reduces then returns those strategy profiles. This is also polynomial time computable.
	\end{enumerate}
Therefore \textsf{PNE-CG} is in $\pls$.
\end{proof}
\begin{definition}{$\pls$-Reductions}{}
	A local search problem $L_1$ is $\pls$-reducible to another local search problem $L_2$, denoted by $L_1\leq_{\pls}L_2$ if there are two polynomial time computable functions $f:D_{L_1}\to D_{L_2}$ and $g:(I_1,s_2)\mapsto s_1$ where $I_1, D_{L_1}$, $s_2\in F_{L_2}(f(I_1))$ and $s_1\in F_{L_1}(I_1)$ such that:\begin{enumerate}[label=(\roman*)]
		\item If $I_1$ is an instance of $L_1$ then $f(I_1)$ is an instance of $L_2$
		\item If $s_2$ is an solution of $f(I_1)$ of $L_2$ then $g(I_1,s_2)$ is a solution for $I_1$ of $L_1$
		\item If $s_2$ is a local optimum for instance $f(I_1)$ of $L_2$ then $f(I_1,s_2)$ has to be a local optimum for instance $I_1$ of $L_1$.
	\end{enumerate}
\end{definition}
%https://panageas.github.io/agt22slides/Lecture9.pdf
\begin{Theorem}{}{}
	The Max Cut Game is $\pls$-complete
\end{Theorem}

We will not prove this but using this theorem we will prove the next theorem about general congestion games.

\begin{Theorem}{}{}
	General Congestion Games are $\pls$-complete
\end{Theorem}
\begin{proof}
	We already shown that general congestion games are in $\pls$. For completeness we will show a $\pls$-reduction from the Max Cut Game. Let $\Gm=(G,w,N)$ be an instance of a max cut game. We will create a congestion game $\Gm'=\Big(E',N, \{S_i\mid i\in[N]\}, \{c_e:[N]\to \bbR_{\geq 0}\mid e\in E\}\Big)$ with $N$ players.
	
	\begin{itemize}
		\item $E'$: For each edge $e\in E$ add two elements $e_L$ and $e_R$ to $E'$. So $E'=\{e_L,e_R\mid e\in E\}$.
		\item $S_i$: Each player $i\in[N]$ has two strategies $S_i=\lt\{\{e_L\mid e\text{ incident on $i$}\}, \{e_R\mid e\text{ incident on $i$}\}\rt\}$. Thus if $e=\{i,j\}\in E$ then the elements $e_L$, $e_R$ can be used by exactly 2 players $i$ and $j$ in $\Gm'$.
		\item $c_e:$ Define $c_{e_L}(1)=c_{e_R}(1)=0$ and $c_{e_L}(2)=c_{e_R}(2)=w(e)$.
	\end{itemize}
This defines a general congestion games. Now consider a strategy profile $s$ in $\Gm'$. Then $C_i(s)=\sum\limits_{\substack{j:\{i,j\}\in E\\ s_j\neq s_i}}e_{i,j}$. Therefore the local minimum in the $\Gm'$ gives a local minimum in $\Gm$. Hence max cut game is $\pls$-reducible to general congestion game.
\end{proof}