\chapter{Introduction}
\section{Matroids}
\dfn{Matroid}{A matroid $M=(E,\mcI)$ has a ground set $E$ and a collection $I$ of subsets of $E$ called the \textit{Independent Sets} st\begin{enumerate}
		\item Downward Closure: If $Y\in \mcI$ then $\forall \ X\subseteq Y$, $X\in \mcI$.
		\item Extension Property: If $X,Y\in \mcI$, $|X|<|Y|$ then $\exs\ e\in Y-X$ such that $X\cup \{e\}$ also written as $X+e\in \mcI$
\end{enumerate}}
\begin{observation*}
	A maximal independent set in a matroid is also a maximum independent set. All maximal independent sets have the same size.
\end{observation*}
\parinf

\textbf{Base:} Maximal Independent sets are called bases.

\textbf{Rank of $\boldsymbol{S\in I}$:} We define the rank function of a matroid $r:\mcP(E)\to \bbZ$ where $r(S)=\max\{|X|\colon X\subseteq S, X\in I\}$ We def

\textbf{Rank of a Matroid:} Size of the base.

\textbf{Span of $\boldsymbol{S\in I}$:} $\{e\in E\colon rank(S)=rank(S+e)\}$

\section{Examples of Matroids}
\subsection{{Uniform Matroid} }It is {denoted as $U_{k,n}$ where $E=[n]$ and $I=\{X\subseteq E\mid |X|\leq k\}$.}


\textbf{Free Matroid: }When $k=n$ we take all possible subsets of $E$ into $I$. This matroid is called {Free Matroid} i.e. $U_{n,n}$\parinn

\subsection{Partition Matroid} Given $E=E_1\sqcup E_2\sqcup \cdots \sqcup E_l$ where $\{E_1,\dots, E_l\}$ is a partition of $E$ and $k_1,\dots,k_l\in \bbN\cup \{0\}$ $$I=\{X\subseteq E\colon |X\cap E_i|\leq k_i\ \forall \ i\in[l]\}$$then $M=(E,I)$ is a partition matroid.
\begin{note}
	If the $E_i$'s are not a partition then suppose $E_1$, $E_2$ has nonempty partition then we will not have a matroid. 
	
	For example: $E_1=\{1,2\}$, $E_2=\{2,3\}$ and $k_1=k_2=1$ then $X=\{1,3\}$ is independent but $Y=\{2\}\subsetneq X$ is not a matroid. 
\end{note}

\subsection{Linear Matroid} Given a $m\times n$ matrix denote its columns as $A_1,\dots, A_n$. Then $$I=\{ X\subseteq [n]\colon \text{Columns corresponding to $X$ are linearly independent}  \}$$\parinn

Here if the underlying field is $\bbF_2$ then it is called \textit{Binary Matroid} and for $\bbF_3$ it is called \textit{Ternary Matroid}.

\subsection{Representable Matroid} A matroid with which we can associate a linear matroid is called a representable matroid.

Eg: $U_{2,3}$. It can be represented by the matrix $A=\mat{1&1&0\\ 1& 0 & 1\\ 0& 1& 1}$, over $\bbF_2$. Over $\bbF_3$ it is same as $U_{3,3}$. 
\nt{There are matroids which are not representable as linear matroids in some field. There are matroids which are not representable on any field as well.}
\lem{}{$U_{2,4}$ is not representable over $\bbF_2$ but representable over $\bbF_3$}

\subsection{Regular Matroid} There are the matroids which are representable over all fields.
\lem{}{Regular Matroids are precisely those which can be represented over $\bbR$ by a Totally Uni-modular matrix}

\subsection{Graphic Matroid / Cyclic Matroid} For a graph $G=(V,E)$ the graphic matroid $M_G=(E,I)$ where $$I=\{F\subseteq E\colon \text{$F$ is acyclic}\}$$

Hence $I$ is the collection of forests of $G$. It follows the downward closure trivially. For extension property let $k=|F_1|<|F_2|=l$ and then there are $n-k$ and $n-l$ components. So $n-k>n-l$. So $\exs$ an edge in $F_2$ which joins 2 components in $F_1$.
\lem{}{A subset of columns is linearly independent iff the corresponding edges don't contain a cycle in the incidence matrix}
\lem{}{Graphic Matroids are Regular Matroids}
\begin{proof-idea}
	Use Incidence Matrix.
\end{proof-idea}
\subsection{Matching Matroids}
We can try to define it like this but it will not work:
\pr{}{Is the following a matroid:
	$E=$ Edges of a graph and $I=\{F\subseteq E\colon \text{$F$ is a matching}\}$
}\parinf

\solve{It is not a matroid since maximal matchings can not be extended to a maximum matching.}
Correct way will be: For a graph $G=(V,E)$ the ground set $=V$ and $$I=\{S\subseteq V\colon \exs \text{a matching that matches all vertices in $S$}\}$$\parinn

The downward closure property trivially holds. For extension property is $|S|<|S'|$ then there exists another vertex in $S'$ which is not matched with $S$, so we can add that vertex to $S$. 


\section{Axiom systems for a matroi}

\subsection{Circuits}
Assume we have a matroid $M=(E,I)$. 
\dfn{Circuit}{A minimal dependent set $C$ such that $\forall\ e\in C$, $C-e$ is an independent set.}
\thm{}{Let $S\in I$. $S+e\notin I$. Then $\exs!$ $C\subseteq S+e$.}
\begin{proof}
	Given $S+e\notin I$. Take the set $\Sg$ where $T\in \Sg$ if $t\notin I$ and $T\subseteq S+e$. $\Sg$ is nonempty since $S+e\in \Sg$. Now under the ordering of inclusion $T$ has a minimal element. Hence this minimal element is the desired circuit $C$ which is minimal dependent set contained in $S+e$.
	
	Now suppose it is not unique. Let $C_1,C_2\subseteq S+e$ be circuits. Suppose $f\in C_1-C_2$. Then $S-e+f$ will still be dependent since $C_2\subseteq S-e+f$. Now by definition we get that $C_1-f$ is independent. Therefore we extend $C_1-f$ to an independent set by adding the elements of $S$ till we reach same size as $|S|$. Now $e\in C$ since $C_1$ was formed because of addition of $e$. Hence if we extend $C_1-f$ till same cardinality as $S$ we will add all the edges of $S$ not in $C_1-f$ except $f$ since adding $f$ will make $C$ be a dependent subset of an independent set which is not possible. Hence $C_1-f$ will be extended to $S-f+e$. Therefore $S+e-f$ is independent which contradicts our previous conclusion that $S+e-f$ is dependent. Hence contradiction.
\end{proof}