\chapter{Polynomial Arithmetic}
\section{Multiplication}
\section{Fast Division}
\begin{algoprob}
	\problemtitle{Polynomial Division}
	\probleminput{$f,g\in\bbF[X]$, $\deg (f,g)\leq d$}
	\problemoutput{Quotient and reminder when $f$ is divided by $g$.}
\end{algoprob}

Suppose $\deg f=a$ and $\deg g=b$. Let $(q,r)\in \bbF[X]$ are the quotient and remainder when $f$ is divided by $g$ i.e. $f=qg+r$. Therefore, $\deg q=a-b$ and $m\coloneqq \deg r<b$. 
%\subsection{Long Division Algorithm}

We can follow the long division algorithm to find $(q,r)$. This algorithm takes $O(a-b)=O(d)$ many iterations to find $q$. And in each iteration we subtract a polynomial from another polynomial by multiplying one of them with power of $x$. For the multiplying with power $x$ is just shifting of the coefficients. For the subtraction of polynomials it takes $O(d)$ time. Therefore, each iteration of the algorithm takes $O(d)$ time complexity. Therefore, the long division algorithm takes $O(d^2)$ time complexity. 

%\subsection{Fast Division Algorithm}
If we can obtain $q$ from $f,g$ then we can get $r$ by following the equation $r=f-gq$.  
\subsection{Reversal of Polynomials}
\begin{idea*}
	Reversal of Polynomials i.e. if $f\in \bbF[X]$ such that $f=f_0+f_1X+\cdots+f_aX^a$ then $$rev(f)=f_0X^a+f_1X^{a-1}+\cdots+f_a=f\lt(\frac1X\rt)X^a$$
\end{idea*}

\nt{We have $\deg f\geq \deg(rev(f))$. Degree of $rev(f)$ can be strictly lesser than the degree of $f$. For example if $f_0=0$ and $f_1\neq 0$, since $rev(f)=X^af\lt(\frac1X\rt)$ the degree of $rev(f)$ is $a-1$.}

So using reversal we will review the equation $f=gq+r$:
\begin{align*}
	& f=qg+r\\
	\iff & X^af\lt(\frac1X\rt)=X^a\lt[q\lt(\frac1X\rt)g\lt(\frac1X\rt)+r\lt(\frac1X\rt)\rt]\\
	\iff & X^af\lt(\frac1X\rt)=cdX^aq\lt(\frac1X\rt)g\lt(\frac1X\rt)+X^ar\lt(\frac1X\rt)\\
	\iff & rev(f)=rev(q)rev(g)+X^{a-m}rev(r) 
\end{align*}
Now we know $a\geq b>m\implies a-m\geq b-m>0$. Therefore, $X^{a-m}rev(r)$ is multiple of some nontrivial power of $X$. Now also we have $$a-m>a-b=\deg q\geq \deg(rev(q))$$ Therefore we have $$rev(f)\equiv rev(q)rev(g)\bmod {X^{a-m}}$$Since $a-m\geq a-b+1$ we have $$rev(q)\bmod {X^{a-m}}\equiv rev(q)\bmod {X^{a-b+1}}\equiv rev(q)$$Therefore we have $$rev(f)\equiv rev(q)rev(g)\bmod {X^{a-b+1}}$$

Hence, it suffices to recover $rev(q)$ in order to recover $q$ from here. So the problem now reduced to finding a solution $h\in\bbF[X]$ for the system $\tdf-h\tdg\equiv 0\bmod{X^N}$.
\subsection{Find solution of \texorpdfstring{$\tdf-h\tdg\equiv 0\bmod {X^N}$}{f-fg=0 mod X\^N}}
\begin{algoprob}
	\problemtitle{Solve $\tdf-h\tdg\equiv 0\bmod {X^N}$}
	\probleminput{$\tdf,\tdg\in\bbF[X]$, $\deg (f,g)\leq d$, $\tdf(0),\tdg(0)\neq 0$ with $N\in\bbN$}
	\problemoutput{Find solution $h$ for the equation $\tdf-h\tdg\equiv 0\bmod {X^N}$}
\end{algoprob}
\begin{lemma}{}{}
	There is an unique $h\in\bbF[X]$ satisfying $\tdf-h\tdg\equiv 0\bmod{X^N}$.
\end{lemma}
\begin{proof}
	Let $\deg \tdf=k$ and $\deg\tdg=l$. Then Suppose $\tdf=\sum\limits_{i=0}^k\tdf_iX^i$ and $\tdg=\sum\limits_{i=0}^l\tdg_iX^i$. Then we can write the equation  $\tdf-h\tdg\equiv 0\bmod{X^N}$ as a linear system like the following: $$\mat{\tdg_0 & \\ \tdg_1 & \tdg_0&\\ \tdg_2&\tdg_1&\tdg_0 & \\ \vdots
		& & & \ddots & \\ & & & & &}\mat{h_0\\ h_1 \\ h_2\\ \vdots \\ h_{k-l}}=\mat{\tdf_0\\ \tdf_1\\ \vdots \\ \tdf_k}$$Lets call the matrix $G$. Since $\tdg_0\neq 0$ the $G$ has nonzero elements in the diagonal. Since the $G$ is lower triangular the determinant of the $G$ is nonzero. Therefore, there exists unique solution for $h$.
\end{proof}

But we don't know how to find inverse of $G$ in near linear time. So we cannot find $h$ like this.
\begin{idea*}
	Find a power series solution for $h=\frac{\tdf}{\tdg}\bmod {X^N}$ in $\bbF\llbracket X\rrbracket\supseteq \bbF[X]$ since in $\bbF\llbracket X\rrbracket$ inverse of $\tdg$ exists
\end{idea*}
\begin{lemma}{}{}
	For every power series $P=\sum\limits_{i=0}^{\infty}P_iX^i\in\bbF\llbracket X\rrbracket$, $P$ has a multiplicative inverse iff $P_0\neq 0$.
\end{lemma}

Since we are dealing with the equation $\tdf-h\tdg\equiv 0\bmod{X^N}$ and $\tdg(0)\neq 0$ there exists a power series solution for $h$. We will see two algorithms to find $h\in \bbF\llbracket X\rrbracket$.

\subsubsection{Algorithm I}
 \begin{align*}
 	\frac{\tdf(X)}{\tdg(X)}\bmod{X^N}  = \frac{\tdf(X)}{\tdg(X)}\;\frac{\tdg(-X)}{\tdg(-X)}\bmod{X^N} = \frac{\tdf(X)\tdg(-X)}{\tdg(X)\tdg(-X)}\bmod{X^N}
 \end{align*}Now $\tdg(X)\tdg(-X)$ is an even function. Therefore, $\exs\ G\in\bbF[X]$ and $\deg G\leq d$ such that $G(X^2)=\tdg(X)\tdg(-X)$. Now we can also decompose $\tdf(X)=\tdf_0(X^2)+X\tdf_1(X^2)$ and $\tdg(-X)=\tdg_0(X^2)+X\tdg_0(X^2)$. Then we have 
\begin{align*}
	 \tdf(X)\tdg(-X) & = \lt(\tdf_0(X^2)+X\tdf_1(X^2)\rt)\lt(\tdg_0(X^2)+X\tdg_1(X^2)\rt)\\
	 & = \underbrace{\lt[ \tdf_0(X^2)\tdg_0(X^2)+X^2\tdf_1(X^2)\tdg_1(X^2)\rt]}_{F_0(X^2)}+X\underbrace{\lt[\tdf_1(X^2)\tdg_0(X^2)+\tdf_0(X^2)\tdg_1(X^2)\rt]}_{F_1(X^2)}\\
	 &=F_0(X^2)+XF_1(X^2) & [\deg F_i\leq d\ \forall i\in\{0,1\}]
\end{align*}Therefore we have \begin{align*}
\frac{\tdf(X)}{\tdg(X)}\bmod{X^N}  & = \frac{F_0(X^2)}{G(X^2)}+X\frac{F_1(X^2)}{G(X^2)}\bmod{X^N} \\
	& = \underbrace{\frac{F_0(X^2)}{G(X^2)}\bmod{X^N}}_{\lt.\frac{F_0(Z)}{G(Z)}\bmod{Z^{\frac{N}2}}\rt|_{Z=X^2}}+X\;\underbrace{\frac{F_0(X^2)}{G(X^2)}\bmod{X^N}}_{\lt.\frac{F_1(Z)}{G(Z)}\bmod{Z^{\frac{N}2}}\rt|_{Z=X^2}}
\end{align*}

Now we recurse. So the algorithm is
\begin{algorithm}
	\DontPrintSemicolon
	\KwIn{$\tdf,\tdg\in\bbF[X]$, $\deg (f,g)\leq d$, $\tdf(0),\tdg(0)\neq 0$ with $N\in\bbN$}
	\KwOut{Find solution $h$ for the equation $\tdf-h\tdg\equiv 0\bmod {X^N}$}
	\Begin{Construct $G,F_0,F_1$\;
	Compute $\frac{F_0(Z)}{G(Z)}\bmod{Z^{\frac{N}2}}$, $\frac{F_1(Z)}{G(Z)}\bmod{Z^{\frac{N}2}}$\;
		Set $Z\longleftarrow X^2$ and combine and return}
	\caption{Solve $\tdf-h\tdg\equiv 0\bmod {X^N}$}
\end{algorithm}

Now we have that $$\frac{F_1(Z)}{G(Z)}\bmod{Z^{\frac{N}2}}=\frac{F_1(Z)\bmod{Z^{\frac{N}2}}}{G(Z)\bmod{Z^{\frac{N}2}}}\bmod{X^{\frac{N}2}}$$Hence the degree got reduced by half. So in the recursion step we can reduce the degree with this. \parinf\vspace{5mm}

\textbf{Time Complexity:} If $T(N)$ is the total time taken while solving for modulo $X^{N}$ then we have the recursion relation $$T(N)\leq 2T\lt(\frac{N}2\rt)+10M(N)$$Hence the total running time of this algorithm is $T(N)=M(N)\log N=Npoly(\log N)$\parinn
\subsubsection{Algorithm II}
Here we can divide $h$ into two parts with each part of degrees $< \frac{N}2$. Then $h(X)=h_0(X)+X^{\frac{N}2}h_1(X)$ where $\deg h_0<\frac{N}2$ and $\deg h_1<\frac{N}2$. Then we have 
$$
	\tdf-\lt(h_0+h_1X^{\frac{N}2}\rt)\tdg\equiv 0\bmod{X^N} \implies (\tdf-h_0\tdg)-X^{\frac{N}2}h_1\tdg\equiv 0 \bmod{X^N} $$Hence we have $\tdf-h_0\tdg\equiv 0\bmod{X^{\frac{N}2}}$. Therefore, we have $$X^{\frac{N}2}\mid \tdf-h_0\tdg\implies \tdf-h_0\tdg=X^{\frac{N}2}$$ Hence $$X^{\frac{N}2}p-X^{\frac{N}2}h_1\tdg\equiv 0\bmod{X^{N}}\implies p-h_1\tdg\equiv 0\bmod{X^{\frac{N}2}}$$
	
	\newpage Therefore we have the following algorithm

\begin{algorithm}
	\DontPrintSemicolon
	\KwIn{$\tdf,\tdg\in\bbF[X]$, $\deg (f,g)\leq d$, $\tdf(0),\tdg(0)\neq 0$ with $N\in\bbN$}
	\KwOut{Find solution $h$ for the equation $\tdf-h\tdg\equiv 0\bmod {X^N}$}
	\Begin{Construct $h_0,h_1$\;
	Solve $\tdf-h_0\tdg\equiv 0\bmod{X^{\frac{N}2}}$\;
$R\longleftarrow \frac{\tdf-h_0\tdg}{X^{\frac{N}2}}$\;
Solve $R-h_1\tdg\equiv 0\bmod{X^{\frac{N}2}}$\;
Output $h_0+X^{\frac{N}2}h_1$
}
	\caption{Solve $\tdf-h\tdg\equiv 0\bmod {X^N}$}
\end{algorithm}\parinf\vspace{5mm}

\textbf{Time Complexity:} If $T(N)$ is the total time taken while solving for modulo $X^{N}$ then we have the recursion relation $$T(N)\leq 2T\lt(\frac{N}2\rt)+O(M(N))$$Hence the total running time of this algorithm is $T(N)=O(M(N)\log N)=Npoly(\log N)$\parinn


\section{Chinese Remainder Theorem}

\section{Derivatives}